---
title: "MedicareLocal_GPPopulation"
author: "DavidFong"
date: "12/03/2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# General Practitioners and population by Medicare Primary Health Network area

```{r}

### start necessary libraries

# data import libraries
library(xlsx)
library(foreign)
library(rgdal)

library(tidyverse)
library(dplyr)
library(lubridate)
# library(rlist)
# library(gdata)

# reproducible research library
library(knitr)

# plotting and maps libraries
library(ggplot2)
library(leaflet)
library(ggmap)

### set figure save path to figure/
knitr::opts_chunk$set(fig.path = 'figure/')
knitr::opts_chunk$set(fig.width=12, fig.height=8) 

```

# Download data files

Population data in Australia 2009 to 2014, grouped by Statistical Area Level 3 (SA3) from http://www.health.gov.au/internet/main/publishing.nsf/Content/PHN-Demographic_Data

Primary Health Network ('PHN') map data from http://www.health.gov.au/internet/main/publishing.nsf/content/phn-boundaries

Concordance table between PHN geographical area and Statistical Area Level 3 from http://www.health.gov.au/internet/main/publishing.nsf/Content/PHN-Concordances

General Practice ('GP', also known as family medicine) workforce by PHN found from Australian Government Health Workforce Data website (http://hwd.health.gov.au/), direct links for General Practitioner data by PHN is not available without login and setting search terms. Pre-downloaded data

```{r pressure, echo=FALSE}
### download data ZIP file if necessary, and extract if necessary

url.phnboundaries <- 'http://www.health.gov.au/internet/main/publishing.nsf/Content/20FD74ABB14A1297CA257F150001FD3B/$File/PHN_boundaries_AUS_May2017_V7_Shapefile.zip'

if (!file.exists('PHN_boundaries_AUS_May2017_V7.shp')) {
  if (!file.exists('PHNboundaries.zip')) {
    if (download.file(url.phnboundaries,'PHNboundaries.zip')) {
      stop('Unable to download data file')
    }
    unzip('PHNboundaries.zip')
  }
}

oz.phn.map <- readOGR('PHN_boundaries_AUS_May2017_V7.shp')
oz.phn.description <- read.dbf('PHN_boundaries_AUS_May2017_V7.dbf')

url.populations <- 'http://www.health.gov.au/internet/main/publishing.nsf/Content/AEADA0341B987748CA257F15000413F6/$File/PopulationReport.xlsx'

if (!file.exists('PopulationReport.xlsx')) {
  if (download.file(url.populations,'PopulationReport.xlsx', method = 'curl')) { # method 'curl' required for Windows
    stop('Unable to download data file')
  }
}

population.2014 <- as.tibble(read.xlsx('PopulationReport.xlsx', sheetName = 'Pop2014', startRow = 9, endRow = 359, header = FALSE))

names(population.2014) <- c('Region','Region2','Nothing','AllAges','Age0_4','Age5_9','Age10_14','Age15_19','Age20_24','Age25_29','Age30_34','Age35_39','Age40_44','Age45_49','Age50_54','Age55_59','Age60_64','Age65_69','Age70_74','Age75_79','Age80_84','Age85plus')

population.2014 <- population.2014 %>%
  select(-Nothing,-Region) %>%
  rename(SA3_NAME_2011 = Region2)
# the original 'Region' variable had leading spaces/tabs

url.concordance <- 'http://www.health.gov.au/internet/main/publishing.nsf/Content/4FAAB33C3ED3E520CA257F150001FD3A/$File/CG_SA3_2011_PHN_2015.xls'
# this is the 2011 SA3 names, which is required for 2009 to 2014 population data

if (!file.exists('CGSA3_2011PHN2015.xlsx')) {
  if (download.file(url.concordance,'CGSA3_2011PHN2015.xlsx', method = 'curl')) {
    stop('Unable to download data file')
  }
}

sa3.phn.concordance <- as.tibble(read.xlsx('CGSA3_2011PHN2015.xlsx', sheetName = 'Table 3', startRow = 8, header = FALSE))

names(sa3.phn.concordance) <-  c('SA3_CODE_2011','SA3_NAME_2011','PHN_CODE','PHN_NAME','RATIO','PERCENTAGE')

gp.numbers <- as.tibble(read.csv('table_2018-03-12_14-41-09.csv', skip = 11, row.names = NULL))
# number of general practitioners 'GP' in each primary health network (PHN) 2016
names(gp.numbers) <- c('PHN_NAME','gp','X')
gp.numbers <- gp.numbers %>%
  select(-X)
gp.numbers[gp.numbers$PHN_NAME == 'Central Queensland','PHN_NAME'] <- 'Central Queensland, Wide Bay, Sunshine Coast'

```
``` {r}
select.age.population <- select(population.2014, SA3_NAME_2011, AllAges)

sa3.phn.concordance <- sa3.phn.concordance %>%
  merge(select.age.population, by = 'SA3_NAME_2011') %>%
  mutate(PHN_Number = as.numeric(AllAges)*as.numeric(RATIO))

phn.population <- sa3.phn.concordance %>%
  group_by(PHN_NAME) %>%
  summarize(patients = sum(PHN_Number))

```

``` {r}

combined.oz.phn.description <- oz.phn.description %>%
  merge(gp.numbers, by = 'PHN_NAME') %>%
  merge(phn.population, by = 'PHN_NAME') %>%
  mutate(pop.gp.ratio = patients/gp)

```

``` {r}
p1 <- leaflet(oz.phn.map) %>% 
  addTiles() %>%
  addPolygons(color="444444",weight=1,
              smoothFactor=0.5,opacity=1.0,fillOpacity = 0.5,
              highlightOptions =
                highlightOptions(color='white',weight=2,bringToFront = TRUE))
```